
In order to take advantages of computer security offered by modern 
cryptography, 
cryptosystems must be realized by cryptographic programs where mathematical
constructs are required to compute on the underlying algebraic
structures of cryptosystems.
Such mathematical constructs are frequently invoked in cryptographic
programs; they are often written in low-level assembly languages and
manually optimized for efficiency. 
% Cryptographic primitives in cryptosystems are but sequences of
% algebraic operations on mathematical structures. 
Security of cryptosystems could be compromised should programming
errors in mathematical constructs be exploited by adversaries.
Subsequently, security guarantees of cryptographic programs
depend heavily on the correctness of mathematical constructs.
%It is therefore of utmost importance to ensure the correctness of
%mathematical constructs used in cryptographic programs. 
In order to build secure cryptosystems, we develop a certified
technique to verify low-level mathematical constructs used in the
security protocol X25519 automatically in this paper.

X25519 is an Elliptic Curve Diffie-Hellman (ECDH) key exchange
protocol; it is a high-performance cryptosystem designed to 
use the secure elliptic curve Curve25519~\cite{Ber06}. Curve25519 is an elliptic
curve offering 128 bits of security when used with ECDH. In addition
to allowing high-speed elliptic curve arithmetic, it is easier to
implement properly, not covered by any known patents, and moreover
less susceptible to implementation pitfalls such as weak 
random-number generators. Its parameters were also selected by
easily described mathematical principles.
% without resorting to any random numbers or seeds. 
These characteristics make Curve25519 a
preferred choice for those who are leery of curves which might have
intentionally inserted
backdoors, such as those standardized by the United States National
Institute of Standards and Technology (NIST). 
Indeed, Curve25519 is currently the
de facto alternative to the NIST P-256 curve. Consequently, X25519 has
a wide variety of applications including the default key exchange
protocol in \openssh since 2014~\cite{W:17:C}.

Most of the computation in X25519, in trade parlance, is in a
``variable base point multiplication,'' and the centerpiece 
is the Montgomery Ladderstep. This is usually a
large constant-time assembly program performing the
finite-field arithmetic that implements the mathematics on Curve25519.
Should the implementation of Montgomery Ladderstep be incorrect, so
would that of X25519. Obviously for all its virtues, X25519 would be
pointless if its implementation is incorrect. This may be even more
relevant in cryptography than most of engineering, because cryptography is
one of the few disciplines with the concept of an omnipresent
adversary, constantly looking for the smallest edge --- and hence
eager to trigger any unlikely event. Revising a cryptosystem
due to rare failures potentially leading to a cryptanalysis is not
unheard of~\cite{HNPPSSW:03:IDFSNE}.
Thus, it is important for security that we can show the computations
comprising the Montgomery Ladderstep or (even better) the X25519
protocol to be correct. 
% However, such verification is not easy due the size both of the
% numbers in play (255 bits and more) and of the program itself
% (10,000+ machine instructions).

Several obstacles need be overcome for the verification of mathematical
constructs in X25519. The key exchange protocol is based on a
group induced by Curve25519. The elliptic
curve is in turn defined over the Galois field $\bbfGF(2^{255} - 19)$. 
To compute on the elliptic curve group, arithmetic computation over
$\bbfGF(2^{255} - 19)$ needs to be correctly implemented. 
Particularly, 255-bit multiplications modulo
$2^{255} - 19$ must be verified. Worse, commodity computing devices do
not support 255-bit arithmetic computation directly. Arithmetic over
the Galois field needs to be implemented by sequences of 32- or 64-bit
instructions of the underlying architectures. One has to
verify that a sequence of 32- or 64-bit instructions indeed
computes, say, a 255-bit multiplication over the finite field. Yet this
is only a single step in the operation on the elliptic curve group.
In order to compute the group operation, another sequence of
arithmetic computation over $\bbfGF(2^{255} - 19)$ is
needed. Particularly, a crucial step,
the Montgomery Ladderstep, requires 18 arithmetic 
computations over $\bbfGF(2^{255} - 19)$~\cite{M:87:SPEC}. 
The entire Ladderstep must be
verified to ensure security guarantees offered by Curve25519.
% and hence the ECDH key exchange protocol X25519.

In this paper, we focus on algebraic properties about low-level
implementations of mathematical constructs in cryptographic programs
as well as range properties about program outputs.
Mathematical constructs by their nature perform computation on
underlying algebraic structures. We aim to verify whether they perform
intended algebraic computation correctly. To this end, we propose the
domain specific language \bvdsl with operations on fixed-width bit-vectors for low-level
mathematical constructs. Algebraic pre- and post-conditions of
programs together with range information about inputs and outputs in \bvdsl are specified as Hoare
triples~\cite{H:69:ABCP}.
Such a specification is converted to static single
assignment form and then translated into (1) an algebraic problem (called
the modular polynomial equation entailment
problem)~\cite{AWZ:88:DQVP,H:07:AENTP} via \zdsl with operations on $\bbfZ$, (2) a range problem, and (3) the absence of program overflows (and underflows). We use the computer
algebra system \singular to solve the algebraic problem~\cite{GP:08:SICA}.
The proof assistant \coq is used to certify the
correctness of translations, as well as solutions to algebraic
problems computed by \singular~\cite{YC:2004:ITPPDC}.
As range problems are hard to be solved automatically with proof assistants, the range problem and the absence of program overflows are verified by SMT (Satisfiability Modulo Theories) solvers.
Correctness of the translation to SMT formulas is again certified by
\coq. The results of SMT solvers however are not certified in our
implementation. A fully certified integration of SMT solvers in \coq
can be used to establish the missing in the future~\cite{EMT+:17:SPISS}.


\begin{figure}
\begin{tikzpicture}[
node distance=0.6cm,
box/.style={rectangle, text width=6cm, fill=white, draw=black},
dot/.style={rectangle, dashed, text width=6cm, fill=white, draw=black},
coord/.style={coordinate},
align=center
]
\node (bvdsl) [box] {algebraic and range specifications in \bvdsl};
\node (bvssa) [box] [below=of bvdsl] {SSA form of algebraic and range specification in \bvdsl};
\node (zssa) [box] [below=of bvssa] {SSA form of algebraic specification in \zdsl};
\node (poly) [box] [below=of zssa] {polynomial equation entailment};
\node (singular) [box] [below=of poly] {\singular};
\node (range) [box] [below=of singular] {range specification};
\node (overflow) [box] [below=of range] {overflow check};
\node (smt) [dot] [below=of overflow] {SMT solvers};
\node (c1) [coord] [left=of bvssa] {};
\node (c2) [coord] [right=of range] {};
\node (c3) [coord] [right=of overflow] {};
\draw[->] (bvdsl) -- node [xshift=1.4cm] {to SSA (Sec.~\ref{subsection:translation:static-single-assignment})} (bvssa);
\draw[-] (bvssa.west) -- node [yshift=-0.7cm,xshift=1.6cm] {to \zdsl (Sec.~\ref{subsection:domain-specific-language-zdsl})} (c1);
\draw[->] (c1) |- node {} (zssa.west);
\draw[->] (c1) |- node {} (range.west);
\draw[->] (c1) |- node {} (overflow.west);
\draw[->] (zssa) -- node  [xshift=1.8cm] {to entailment (Sec. \ref{subsection:translation:multivariant-polynomial-equations})} (poly);
\draw[->] (poly) -- node  [xshift=1.4cm] {solved by (Sec.~\ref{subsection:solving-algebraic-equations})} (singular);
\draw[-] (range.east) -- node {} (c2);
\draw[-] (overflow.east) -- node {} (c3);
\draw[->] (c2) |- node {} (smt.east);
\draw[->] (c3) |- node [xshift=-1.4cm,yshift=0.5cm] {solved by (Sec.~\ref{subsection:solving-range-overflow-checks})} (smt.east);
\end{tikzpicture}
\caption{The verification flow. Except the answers from SMT solvers, all the translations and the answers from \singular are certified by \coq.}
\label{fig:overview}
\end{figure}

We report case studies on verifying mathematical constructs used in
the X25519 ECDH key exchange protocol~\cite{BDL+:11:HSHSS,BDL+:12:HSHSS}.
For each arithmetic operations
(such as addition, subtraction, and multiplication) over $\bbfGF(2^{255} - 19)$,
their low-level real-world implementations are converted to our domain
specific language \bvdsl manually. We specify algebraic properties
of mathematical constructs in Hoare triples. Mathematical constructs
are then verified against their algebraic
specifications with our automatic technique.
The implementation of the Montgomery Ladderstep is
verified similarly.

We have the following contributions:
\begin{itemize}
\item We propose a domain specific language \bvdsl for modeling low-level
  mathematical constructs used in cryptographic programs.
\item We give a certified verification condition generator from
  algebraic specifications of programs to the modular polynomial
  equation entailment problem.
\item We give a certified translation from range problems and overflow checks to SMT formulas.
\item We verify arithmetic computation over a finite field of order
  $2^{255} - 19$ and a
  critical program (the Montgomery Ladderstep) automatically.
\item To the best of our knowledge, our work is the first automatic
  and certified verification on real cryptographic programs with
  minimal human intervention.
\end{itemize}

\input{related-work}


This paper is organized as follows. After preliminaries
(Section~\ref{section:preliminaries}), our domain specific
language is described in Section~\ref{section:domain-specific-language-bvdsl}.
Section~\ref{section:translation}
presents the translation to the algebraic
problem. A certified translation from range and overflow checks to SMT formulas plus a certified solver for the algebraic problem are discussed in Section~\ref{section:verification-of-specifications}.
Section~\ref{section:evaluation} contains experimental results. It is
followed by conclusions.

%%% Local Variables:
%%% mode: latex
%%% eval: (TeX-PDF-mode 1)
%%% eval: (TeX-source-correlate-mode 1)
%%% TeX-master: "certified_vcg"
%%% End:
