
We write $\mathbb{B} = \{ \mathit{ff}, \mathit{tt} \}$ for the Boolean
domain. Let $\bbfN$ and $\bbfZ$ denote all natural numbers and all integers
respectively. We use $[n]$ to denote the set $\{ 0, 1, \ldots, n \}$
for $n \in \bbfN$.

A \emph{monoid} $\mathcal{M} = (M, \epsilon, \cdot)$ consists of a set
$M$ and an associative binary operator $\cdot$ on $M$ with the
\emph{identity} $\epsilon \in M$. That is, $\epsilon \cdot m = m \cdot
\epsilon = m$ for every $m \in M$.
A \emph{group} $\mathcal{G} = (G, 0, +)$ is an algebraic structure
where $(G, 0, +)$ is a monoid and there is a $-a \in G$ such that
$(-a) + a = a + (-a) = 0$ for every $a \in G$. The element $-a$ is
called the \emph{inverse} of $a$. $\mathcal{G}$ is \emph{Abelian} if
the operator $+$ is commutative.
A \emph{ring} $\mathcal{R} = (R, 0, 1, +, \times)$ with $0 \neq 1$ is
an algebraic structure such that
\begin{itemize}
\item $(R, 0, +)$ is an Abelian group; 
\item $(R, 1, \times)$ is a monoid; and 
\item $\times$ is distributive over $+$: $a \times (b + c) = a \times
  b + a \times c$ for every $a, b, c \in R$.
\end{itemize}
If $\times$ is commutative, $\mathcal{R}$ is a \emph{commutative}
ring. 
% An \emph{integral domain} is a commutative ring where $a \times
% b = 0$ implies $a = 0$ or $b = 0$. 
A \emph{field} $\mathcal{F}$ $=$ $(F,$
$0,$ $1,$ $+,$ $\times)$ is a commutative ring where $(F\!\setminus\!\{0\}, 1, \times)$ is also
a group. $(\bbfN, 1, \times)$ is a monoid. $(\bbfZ, 0, 1, +, \times)$ 
is a commutative ring but not a field. 
For any prime number $\varrho$, the set $\{ 0, \ldots, \varrho - 1 \}$
with the addition and multiplication modulo $\varrho$ forms a \emph{Galois
field} of order $\varrho$ (written $\bbfGF(\varrho)$).
We focus on Galois fields of very large orders, in particular, $\myprime =
2^{255} - 19$.

Fix a set of variables $\vx$. $\mathcal{R}[\vx]$ is the set of
polynomials over $\vx$ with coefficients in the ring
$\mathcal{R}$. $\mathcal{R}[\vx]$ is a ring. A set $I \subseteq
\mathcal{R}[\vx]$ is an \emph{ideal} if 
\begin{itemize}
\item $f + g \in I$ for every $f, g \in I$; and
\item $h \times f \in I$ for every $h \in
  \mathcal{R}[\vx]$ and $f \in I$. 
\end{itemize}
Given $G \subseteq \mathcal{R}[\vx]$, $\langle G \rangle$ is the
minimal ideal containing $G$; $G$ are the \emph{generators}
of $\langle G \rangle$. The \emph{ideal membership}
problem is to decide if $f \in I$ for a given ideal $I$ and $f
\in \mathcal{R}[\vx]$.

Let $\BV^w$ be the set of all bit-vectors with a bit-width $w$.
The unsigned value of $b \in \BV^w$ is denoted by $\unsigned{b}$.
For a natural number or an integer $n$, let $\bv^w(n)$ be the two's complement representation of $n$ in a bit-width $w$.
We use the following common operators for fixed-width bit-vectors: $\BV^w +_\bvop \BV^w : \BV^w$ for addition, $\BV^w -_\bvop \BV^w : \BV^w$ for subtraction, $\BV^w \times_\bvop \BV^w : \BV^w$ for multiplication, $\BV^{w_1} ._\bvop \BV^{w_2} : \BV^{w_1 + w_2}$ for concatenation, $\BV^w \#_\bvop n : \BV^{w+n}$ for zero extension, $\BV^w \ll_\bvop n : \BV^w$ for left-shifting, $\BV^w \gg_\bvop n : \BV^w$ for logical right-shifting, and $\BV^w[i,j] : \BV^{i - j + 1}$ with $0 \leq j \leq i < w$ for bits extraction.
We also assume comparison operators $<_\bvop$ and $\leq_\bvop$ between unsigned values of bit-vectors.

Given a bit-vector $b \in \BV^{2w}$, define $\hi_\bvop(b) \defn b[2w - 1, w]$ for the extraction of higher $w$ bits, and $\lo_\bvop(b) \defn b[w - 1, 0]$ for the extraction of lower $w$ bits.
For operations $\bullet \in \{+_\bvop, -_\bvop, \times_\bvop\}$, we define their extended version $\bullet^\#$ which performs the original operation after doubling the width of operands by zero extension.
In the extended operations, the width of operands is doubled only once.
For example, for $b_1, b_2, b_3 \in \BV^{w}$, we have $b_1$ $+_\bvop^\#$ $b_1$ $\defn$ $(b_1$ $\#_\bvop$ $w)$ $+_\bvop$ $(b_2$ $\#_\bvop$ $w)$ and $b_1$ $+_\bvop^{\#}$ $b_2$ $+_\bvop^{\#}$ $b_3$ $\defn$ $(b_1$ $\#_\bvop$ $w)$ $+_\bvop$ $(b_2$ $\#_\bvop$ $w)$ $+_\bvop$ $(b_3$ $\#_\bvop$ $w)$.

%Define six more operations which takes high bits or low bits after $\bullet^{\#}$ operation.
%\[
%\begin{array}{rcl}
%b_1 \bullet^{\hi} b_2 & \defn & \hi_\bvop(b_1 \bullet^{\#} b_2) \\
%b_1 \bullet^{\lo} b_2 & \defn & \lo_\bvop(b_1 \bullet^{\#} b_2) \\
%\end{array}
%\]


%%% Local Variables:
%%% mode: latex
%%% eval: (TeX-PDF-mode 1)
%%% eval: (TeX-source-correlate-mode 1)
%%% TeX-master: "certified_vcg"
%%% End:
