
Given $q_a, q_a' \in \bvPreda$, $q_r, q_r' \in \bvPredr$, and $p \in \bvProg$, we reduce the problem of checking $\models \hoaretriple{q_a \andar q_r}{p}{q_a' \andar q_r'}$ to (1) the entailment problem of modular
polynomial equations over integer variables proving $\models \hoaretriple{q_a}{p}{q_a'}$ via an intermediate language \zdsl, (2) a range problem $\models \hoaretriple{q_r}{p}{q_r'}$, and (3) a safety check of program $p$.
The reduction is carried out by the following three transformations:
\begin{enumerate}
%\item Program slicing. To improve efficiency, fragments of the program $p$
%  irrelevant to the post-condition $q'$ are removed by program
%  slicing (Section~\ref{subsection:translation:slicing})~\cite{W:81:PS}.
\item Static single assignments. The program is transformed
  into static single assignments. Variables in pre- and
  post-conditions are also renamed
  (Section~\ref{subsection:translation:static-single-assignment})~\cite{AWZ:88:DQVP}.
\item \zdsl. The algebraic specification $\hoaretriple{q_a}{p}{q_a'}$ in \bvdsl is transformed to a specification in \zdsl so that the validity of the specification in \zdsl implies the validity of $\hoaretriple{q_a}{p}{q_a'}$ in \bvdsl if the program $p$ is safe. (Section~\ref{subsection:domain-specific-language-zdsl}).
\item Modular polynomial equations. Validity of algebraic specifications in \zdsl
  is reduced to the entailment of modular polynomial equations
  (Section~\ref{subsection:translation:multivariant-polynomial-equations})~\cite{H:07:AENTP}.
\end{enumerate}

For each transformation, we give an algorithm and establish the
correctness of the algorithm in \coq~\cite{YC:2004:ITPPDC}.
Specifically, semantics for \zdsl and validity of specifications in \zdsl are formalized.
The correctness of transformations is then certified by the proof assistant \coq.
For static single assignments, we
construct machine-checkable proofs for the soundness and completeness
of the transformation. For modular polynomial equations, another
\coq-certified proof shows the soundness of the transformation
from the validity of the algebraic specification to the entailment of
modular polynomial equations. In the following subsections,
transformations and their correctness are elaborated in details.
