
The last step transforms any algebraic program specification to the 
modular polynomial equation entailment problem defined as follows. For
$q \in \Pred$, we write $q(\vx)$ to signify the free variables $\vx$
occurred in $q$. Given $q(\vx), q'(\vx) \in \Pred$, the \emph{modular
  polynomial equation entailment} problem decides whether 
$q(\vx) \implies q'(\vx)$ holds when $\vx$ are substituted for
arbitrary integers. That is, we want to check if $q(\vx) \implies
q'(\vx)$ evaluates to $\btt$ after each variable $x$ is replaced by
$\nu(x)$ for every valuation $\nu \in \St$. We write $\bbfZ \models
\forall \vx. q(\vx) \implies q'(\vx)$ if it is indeed the case.

Programs in static single assignment form can easily be transformed to
conjunctions of polynomial equations. We first show how to transform
statements into polynomial equations. An assignment statement is
translated to a polynomial equation with a single variable on the left
hand side. For a $\dslcode{Split}$ statement, it is transformed to an
equation with a linear expression of the assigned variables on the
left hand side (Algorithm~\ref{algorithm:polynomial-statements}). 
\begin{algorithm}
  \begin{algorithmic}[1]
    \Function{StmtToPolyEQ}{$s$}
    \Match{$s$}
      \Case{$v \leftarrow e$}
        \Return $v = e$
      \EndCase
      \Case{$\concat{v_h}{v_l} \leftarrow \dslcode{Split}(e, n)$}
        \Return $v_l + 2^n v_h = e$
      \EndCase
    \EndMatch
    \EndFunction
  \end{algorithmic}
  \caption{Polynomial Equation Transformation for Statements}
  \label{algorithm:polynomial-statements}
\end{algorithm}

A program in static single assignment form is transformed to the
conjunction of polynomial equations corresponding to its statements
(Algorithm~\ref{algorithm:polynomial-programs}). 

\begin{algorithm}
  \begin{algorithmic}[1]
    \Function{ProgToPolyEQ}{$p$}
    \Match{$p$}
      \Case{$\epsilon$} \Return $\top$ \EndCase
      \Case{$s; pp$}
        \Return \Call{StmtToPolyEQ}{$s$} $\wedge$
                \Call{ProgToPolyEQ}{$pp$}
      \EndCase
    \EndMatch
    \EndFunction
  \end{algorithmic}
  \caption{Polynomial Equation Transformation for Programs}
  \label{algorithm:polynomial-programs}
\end{algorithm}

Algorithm~\ref{algorithm:polynomial-statements}
and~\ref{algorithm:polynomial-programs} are specified straightforwardly 
in \gallina. We use the proof assistant \coq to prove properties
about the algorithms. First, note that $\textsc{ProgToPolyEQ}(p) \in
\Pred$ for every $p \in \Prog$. The following theorem shows that any
behavior of the program $p$ is a solution to the system of polynomial
equations $\textsc{ProgToPolyEQ}(p)$. In other words,
$\textsc{ProgToPolyEQ}(p)$ gives an abstraction of the program $p$.

\begin{theorem}
  Let $p \in \Prog$ be a well-formed program in static single assignment
  form. For every $\nu, \nu' \in \St$ with $\nu \goesto{p} \nu'$, 
  we have $\bbfZ \models \textsc{ProgToPolyEQ}(p) [\nu']$.
\end{theorem}

Definition~\ref{definition:program-entailment} gives the modular
polynomial equation entailment problem corresponding to an algebraic
program specification.
\begin{definition}
  For any $q, q' \in \Pred$ and $p \in \Prog$ in static single assignment
  form, define
  \begin{eqnarray*}
    \Pi(\hoaretriple{q}{p}{q'}) & \defn &
    q(\vx) \wedge \varphi(\vx) \implies q'(\vx)
  \end{eqnarray*}
  where $\varphi(\vx) =
  \textsc{ProgramToPolyEQ}(p)$. 
  \label{definition:program-entailment}
\end{definition}

\begin{figure}
  \centering
  \[
  \begin{array}{l}
  \top \wedge
  \left(
  \begin{array}{lclcl}
    \begin{array}{rclc}
      {r}^0_0 & = & {x}^0_0 & \wedge \\
      {r}^0_1 & = & {x}^0_1 & \wedge \\
      {r}^0_2 & = & {x}^0_2 & \wedge \\
      {r}^0_3 & = & {x}^0_3 & \wedge \\
      {r}^0_4 & = & {x}^0_4 & \wedge \\
    \end{array}
    &\hspace{.01\textwidth}&
    \begin{array}{rclc}
      {r}^1_0 & = & {r}^0_0 + {4503599627370458} & \wedge \\
      {r}^1_1 & = & {r}^0_1 + {4503599627370494} & \wedge \\
      {r}^1_2 & = & {r}^0_2 + {4503599627370494} & \wedge \\
      {r}^1_3 & = & {r}^0_3 + {4503599627370494} & \wedge \\
      {r}^1_4 & = & {r}^0_4 + {4503599627370494} & \wedge\\
    \end{array}
    &\hspace{.01\textwidth}&
    \begin{array}{rclc}
      {r}^2_0 & = & {r}^1_0 - {y}^0_0 & \wedge \\
      {r}^2_1 & = & {r}^1_1 - {y}^0_1 & \wedge \\
      {r}^2_2 & = & {r}^1_2 - {y}^0_2 & \wedge \\
      {r}^2_3 & = & {r}^1_3 - {y}^0_3 & \wedge \\
      {r}^2_4 & = & {r}^1_4 - {y}^0_4
    \end{array}
  \end{array}
  \right)\\
    \hspace{.05\textwidth}
    \implies 
    \mathit{radix}({x}^0_4, {x}^0_3, {x}^0_2, {x}^0_1, {x}^0_0) -
    \mathit{radix51}({y}^0_4, {y}^0_3, {y}^0_2, {y}^0_1, {y}^0_0)
    \equiv
    \mathit{radix51}({r}^2_4, {r}^2_3, {r}^2_2, {r}^2_1, {r}^2_0)
    \mod \myprime
  \end{array}
  \]
  \caption{Modular Polynomial Equation Entailment for \dslcode{subSSA}}
  \label{figure:translation:subtraction-polynomial}
\end{figure}

\noindent
\emph{Example}.
The modular polynomial equation entailment problem coresponding to the
algebraic specification of subtraction is shown in
Figure~\ref{figure:translation:subtraction-polynomial}. The problem
has 15 polynomial equality constraints with 25 integer variables. We would like
to see if $\mathit{radix51}({r}^2_4, {r}^2_3, {r}^2_2,
{r}^2_1, {r}^2_0)$ is the difference between $\mathit{radix}({x}^0_4,
{x}^0_3, {x}^0_2, {x}^0_1, {x}^0_0)$ and $\mathit{radix51}({y}^0_4,
{y}^0_3, {y}^0_2, {y}^0_1, {y}^0_0)$ in $\bbfGF(\varrho)$ under the
constraints. 

Theorem~\ref{theorem:program-to-q-soundness} establishes the
soundnesss of Algorithm~\ref{algorithm:polynomial-programs}. Its proof
is again certified by \coq.
\begin{theorem}
  \label{theorem:program-to-q-soundness}
  Let $q, q' \in \Pred$ be predicates, and $p \in \Prog$ a program in
  static single assignment form and well-formed. 
  If $\bbfZ \models \forall \vx.\Pi(\hoaretriple{q}{p}{q'})$, then
  $\models$ $\hoaretriple{q}{p}{q'}$.
\end{theorem}

