
Given $q, q' \in \Pred$ and $p \in \Prog$, we reduce the problem of checking
$\models \hoaretriple{q}{p}{q'}$ to a problem of checking (modular)
polynomial equalities constrained by a system of polynomial equations over
integers. The reduction is carried out by the following three
transformations:
\begin{enumerate}
\item Program slicing. To improve efficiency, the segments of the program $p$
  irrelevant to the post-condition $q$ are removed by program slicing 
  (Section~\ref{subsection:translation:slicing}). 
\item Static single assignments. The sliced program is transformed
  into static single assignments. Variables in the pre- and
  post-conditions are renamed accordingly
  (Section~\ref{subsection:translation:static-single-assignment}).
\item Polynomial equations. The resultant algebraic
  specification is transformed into checking (modular)
  polynomial equalities constrained by a system of polynomial equations
  (Section~\ref{subsection:translation:multivariant-polynomial-equations}).
\end{enumerate}

For each transformation, we give an algorithm and establish the
correctness of the algorithm in \coq. Specifically, a semantics
for our domain specific language is given and 
$\models \hoaretriple{q}{p}{q'}$ is formalized in \coq. Then the
correctness of transformations is established by formal \coq
proofs. For program slicing and static single assignments, we
construct machine-checkable proofs for the soundness and completeness
of the two transformations. For polynomial equations, another
\coq-certified proof shows the soundness of the transformation
from the validity of the algebraic specification to solvability of an
algebraic problem over (modular) polynomial equations. In the
following subsections, we elaborate the transformations and their
correctness in details. 